---
title: "HW15"
author: '108048110'
date: '2022-05-25'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BACS HW - Week 15

------------------------------------------------------------------------

## Prerequisite

```{r message=FALSE, warning=FALSE}
library(seminr)
```

```{r warning=FALSE}
security_score <- read.csv("data/security_data_sem.csv")
```

------------------------------------------------------------------------

## What is SEM?

-   SEM, structural equation modeling, aka covariance structural modeling, casual modeling, analysis of moment structures.

-   SEM is suitable when the concepts researcher is interested in are complex or multifaceted, it focuses on indirect or mediated, and performs well on modeling casual system.

-   Key to represent SEM:

    $X = t + e$

    -   $X$ is the thing that we want to measure.

    -   $t$ is the true score, the true underlying level of happiness.

    -   $e$ is the error term.

        -   Systematic error: Bias, the mean of the individual errors does not cancelled out.

        -   Random error: $mean(e) = 0$

-   One can thought of SEM as path analysis using latent variables.

    ### Latent Variable

    -   Most social scientific concepts are not directly observable.

        -   e.g. Intellegent, Honesty, Happiness, etc.

    -   This makes them hypothetical or latent.

        -   Latent within people at some level that drive attitude and behavior, but not observable.

        -   Unable to measure it directly.

    -   By using latent variables, we are able to cover the full conceptual map, remove and reduce random errir in `measured construct.`

    ### Path analysis

    -   Diagrammatic representation of a theoretical model using standardized notion.

    -   Regression equations are specified between measured variables.

        -   **Note.** in standard path analysis, we would not use latent variables, instead, the variables that are directly observed.

    -   Standardized notion

        |                                        |                                       |
        |----------------------------------------|---------------------------------------|
        | ellipse                                | measured latent variable              |
        | rectangle                              | observed/manifest variable, indicator |
        | circle                                 | error variance, disturbance term      |
        | Curved, double headed arroes           | covariance/ non-directional path      |
        | Straight arrows point in one direction | regression/ directional path          |

        : standardized notion

------------------------------------------------------------------------

## PLS-SEM

**Partial least squares SEM**

For many years, covariance-based structural equation modeling (CB-SEM) was the dominant method for analyzing complex interrelationships between observed and latent variables. The PLS-SEM is appealing to many researchers as it enables them to estimate complex models with many constructs, indicator variables and structural paths without imposing distributional assumptions (i.e. the dependent variable must be approximately normally distributed, and the relationships among the variables are assumed to be casual, linear and additive) on the data.

-   PLS-SEM is a causal-predictive approach to SEM that emphasizes prediction in estimating statistical models, whose structures are designed to provide casual explanations.

#### When to use PLS-SEM?

The most prominent justifications for using PLS-SEM are attributed to...

-   Non-normal data

    Data collected for social science research often fails to follow a multivariate normal distribution.

    When attempting to evaluate a path model using CB-SEM, non-normal data can lead to underestimated standard errors and inflated goodness-of-fit measures.

-   Small sample sizes

    Different from CB-SEM, PLS-SEM can be utilized with much smaller sample sizes, even when models are highly complex.

    In these situations, PLS-SEM generally achieves higher levels of statistical power and demonstrates much better convergence behavior than CB-SEM.

-   Formative Indicators

    Analyzing formative indicators with CB-SEM often leads to identification problems ([Jarvis *et al.*, 2003](https://www.emerald.com/insight/content/doi/10.1108/EBR-10-2013-0128/full/html#b36){alt=""}). However, formative indicators should be approached with caution when using PLS-SEM.

    Researchers should be aware that the evaluation of formatively measured constructs relies on a totally different set of criteria compared to their reflective counterparts. Researchers should apply the most recent set of evaluation criteria when examining the validity of formatively measured constructs.

------------------------------------------------------------------------

**To sum up...** Depending on the specific empirical context and objectives of the study, PLS-SEM's distinctive methodological features make it a particularly valuable and potentially better-suited alternative to the more popular CB-SEM approaches in practical applications.

Generally, however, neither method is superior to the other overall. Rather, the selection of the proper method depends on the objective of the study.

------------------------------------------------------------------------

## Question 1) Composite Path Models using PLS-PM

### a. Create a PLS path model

#### *i*. Measurement model

-   All constructs are measured as composites.

    ```{r}
    sec_mm <- constructs(
      composite("TRUST", multi_items("TRST", 1:4)),
      composite("SEC", multi_items("PSEC", 1:4)),
      composite("REP", multi_items("PREP", 1:4)),
      composite("INV", multi_items("PINV", 1:3)),
      composite("POL", multi_items("PPSS", 1:3)),
      composite("FAML", single_item("FAML1")),
      interaction_term(iv="REP", moderator="POL", method=orthogonal)
      )
    ```

#### *ii.* Structural Model

-   Paths between constructs

    ```{r}
    sec_sm <- relationships(
      paths(from=c("REP","INV","POL","FAML", "REP*POL"),
            to = "SEC"),
      paths(from="SEC",
            to = "TRUST")
    )
    ```

### b. Showing the result

```{r}
sec_pls <- estimate_pls(
  data = security_score,
  measurement_model = sec_mm,
  structural_model = sec_sm
)
```

```{r}
sec_pls$path_coef

sec_pls_summary <- summary(sec_pls)
```

#### *i.* Plot a figure of the estimated model.

```{r}
plot(sec_pls)
```

#### \* ii.\* Weights and loadings of composites.

```{r}
# weights
sec_pls_summary$weights
```

-   **Adding these weights together gets the scores of the entire construct.**

```{r}
# loadings
sec_pls_summary$loadings
```

-   **The correlation between the composite and the construct.**

```{r}
# VIF
sec_pls_summary$validity$vif_items
```

#### *iii.* Regression coefficients of paths between factors.

```{r}
# where the path fits the R^2
sec_pls_summary$paths
```

```{r}
sec_pls_summary$descriptives$statistics$constructs
sec_pls_summary$descriptives$correlations$constructs
```

#### *iv.* Bootstrapped path coefficients: t-values, 95% CI.

```{r}
boot_sec_pls <- bootstrap_model(
  seminr_model = sec_pls,
  nboot = 1000,
  seed = 42
)
```

```{r warning=FALSE}
# a matrix of the bootstrap path coefficients and standard deviations
boot_sec_pls$paths_descriptives
boot_sec_pls$total_paths_descriptives

boot_sec_pls_summary <- summary(boot_sec_pls)
#reports a matrix of direct paths and their standard deviation, t_values, and confidence intervals.
boot_sec_pls_summary$bootstrapped_paths
#reports a matrix of total paths and their standard deviation, t_values, and confidence intervals.
boot_sec_pls_summary$bootstrapped_total_paths

# evaluate the mean estimate, standard deviation, t_value and confidence intervals for specific paths
# from sec to trust
specific_effect_significance(boot_sec_pls, 
                             from = "SEC",
                             to = "TRUST",
                             alpha = 0.05
                             )

```

------------------------------------------------------------------------

## Question 2) Common-Factor Models using CB-SEM

### a. Create a common factor model

#### *i.* Convert your earlier measurement model to being entirely reflective.

```{r}
sec_mm_reflective <- as.reflective(sec_mm)
```

-   **But we often wish to conduct a CFA of our measurement model prior to CBSEM**.

```{r}
sec_cfa <- estimate_cfa(data=security_score,
                        sec_mm_reflective)
sec_cfa_summary <- summary(sec_cfa)
sec_cfa_summary$descriptives$correlations$constructs
plot(sec_cfa)
```

#### *ii.* Use the same structural model as before.

```{r}
sec_cbsem <- estimate_cbsem(data=security_score,
                            sec_mm_reflective,
                            sec_sm)
sec_cbsem_summary <- summary(sec_cbsem)
sec_cbsem_summary$descriptives$correlations$constructs
```

### b. Showing the result

#### *i.* Plot a figure of the estimated model.

```{r}
# CB-SEM
plot(sec_cbsem)
```

#### *ii.* Show the loadings of composites.

```{r}
sec_cbsem_summary$loadings$significance
```

#### *iii.* Show the regression coefficients of paths between factors, and their p-values.

```{r}
sec_cbsem_summary$paths$significance
```

#### 
